---
title: "5 Parts of Speech"
description: |
  The frequency of various parts of speech.
author:
  - name: Lars Hinrichs
    url: https://larshinrichs.site
    affiliation: The University of Texas at Austin
    affiliation_url: https://liberalarts.utexas.edu/english
date: 12-22-2020
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = TRUE,
                      cache.lazy = FALSE)
library(pacman)
p_load(rio, dplyr, ggplot2, stringr, knitr, kableExtra,
       beepr)
```

## Data tagging

The four speeches will be part-of-speech tagged for this analysis, so that each word will be marked for its grammatical class.

```{r run-pos-tagging, eval=F}
# init tagging
p_load(scriptuRs, cleanNLP, reticulate)
pypath <- "/usr/local/opt/python/bin/python3.7"
use_python(pypath)
cnlp_init_spacy()

srpath <- "../../data/corpus.RDS"

# perform tagging
corpus_for_tagging <- 
  rio::import(srpath) %>% 
  mutate(linelength = str_count(text, pattern = " ")) %>%
  filter(linelength > 2)

corpus_annotated <- cnlp_annotate(corpus_for_tagging,
                                  text_name = "text", 
                                  verbose = T)

corpus_postagged <- 
  corpus_annotated$token %>% 
  left_join(corpus_annotated$document)

rm(corpus_for_tagging)
rm(corpus_annotated)

corpus_postagged <- 
  corpus_postagged %>% 
  select(-linelength) %>% 
  filter(! upos %in% c("PUNCT", "SPACE"))

corpus_postagged %>% rio::export(
  "data/corpus_postagged.RDS")

beepr::beep(sound = 2, expr = NULL)

```

```{r show-data}
corpus_postagged <- 
  rio::import("data/corpus_postagged.RDS")
corpus_postagged %>% 
  head(10) %>% 
  select(-party, -nickname, -first_name) %>% 
  kbl(caption = "Top of data frame, POS-tagged.") %>% 
  kable_paper()
```

```{r get-totals}
n_total <- 
  corpus_postagged %>% 
  group_by(doc_id) %>% 
  summarise(total = n())

n_pos <- 
  corpus_postagged %>% 
  group_by(doc_id) %>% 
  count(upos) %>% 
  left_join(n_total) %>% 
  mutate(freq_per_1000 = n/total*1000)
```

## Comparing POS frequencies

We will work with the `upos` column, which gives a fine enough classification of POS into 14 different types:

```{r}
corpus_postagged %>% pull(upos) %>% unique()
```

The unit of analysis will be frequency per 1,000 words, to make the numbers comparable.

### Global frequencies

The following plot shows these values ordered by POS-tag.

```{r show-freqs-panel, fig.cap="Comparison of POS frequencies.", layout="l-body-outset"}

pal <- wesanderson::wes_palette("Darjeeling2", 4, type = "discrete")
n_pos %>% 
  ggplot(aes(x=doc_id,
             y=freq_per_1000,
             fill=doc_id)) +
  geom_col(width = .45) +
  facet_wrap(~upos, scales = "free") +
  labs(
    title = "Parts of Speech in Texan Inaugurals",
    subtitle = "Frequencies per 1,000 words",
    x = NULL, y = NULL,
       caption = "www.texasenglish.org") +
  scale_fill_manual(values = pal) +
  theme_bw(base_family = "Volkhov-Regular",
                  base_size = 10) +
  theme(axis.text.x = element_blank(),
        legend.title = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.text = element_text(size=7)
        )

```

### Nouns vs. Verbs

A big point of interest is the degree of overall nominal vs. verbal nature of a text. There is a group of parts of speech that pattern together with nouns. They are:

-   noun (NOUN)

-   proper noun (PROPN)

-   adjective (ADJ)

-   determiner (DET)

-   preposition (ADP)

The other group of tags that co-occur and that characterize a more verbal style are

-   verb (VERB)

-   auxiliary (AUX)

-   adverb (ADV)

-   conjunction (CCONJ, SCONJ)

Let us treat the two groups in aggregate and compare the speeches for their frequencies.

```{r nominal-v-verbal, layout="l-body-outset", fig.cap="Aggregate counts for nominal and verbal POS-groups in the speeches.", fig.height=3}
nominal <- c("NOUN", "PROPN", "ADJ", "DET",
             "ADP")
verbal <- c("VERB", "AUX", "ADV", "CCONJ", "SCONJ")

corpus_postagged <- 
  corpus_postagged %>% 
  mutate(posgroup = case_when(
    upos %in% nominal ~ "nominal",
    upos %in% verbal ~ "verbal",
    TRUE ~ "other"
  ))

n_posgroup <- 
  corpus_postagged %>% 
  count(doc_id, posgroup) %>% 
  left_join(n_total) %>% 
  mutate(freq = n/total*100)

n_posgroup %>% 
  filter(posgroup != "other") %>% 
  ggplot(aes(x=posgroup,
             y=freq,
             fill=doc_id)) +
  geom_col(show.legend = F,
           position = "dodge") +
  scale_fill_manual(values=pal) +
  labs(title="Nominal vs. Verbal Style in Texan Inaugurals",
       subtitle="Frequencies per 100 words",
       x=NULL, y=NULL,
       caption="www.texasenglish.org"
       ) +
  facet_wrap(~doc_id, nrow=1) +
  theme_bw(base_family = "Volkhov-Regular",
                  base_size = 10) +
  theme(legend.title = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank())
  
```

The speech that truly stands out (again) is 2005-Bush. He makes a clear switch to more nominal style, compensated by a drop in verbal POSs, relative to his Texan predecessors.

## Interpretation

### What is a "more nominal" style?
